\clearpage 
%\addcontentsline{toc}{chapter}{Literatúra}
\begin{thebibliography} {99}
 %\addcontentsline{toc}{section}{Literatura}

\bibitem{Baker} Baker A., (2006), ``Matrix Groups: An Introduction to Lie Group Theory'', 3rd edition, Princeton Springer Great Britain.  
\bibitem{Bellman} Bellman R.E., (1957), ``Dynamic Programming'', Princeton University Press, NJ. 
\bibitem{Benchmark} Bielicky T.R., Chancelier J. P., Pliska S. R., Sulem A., (2003), ``Risk sensitive portfolio optimization with transaction cost'', {\em Journal of Computational Finance, 8(1), 39-63}. 
\bibitem{Benchmark2} Bielicky T.R., Pliska S. R., (2000), ``Risk sensitive dynamic asset management in the presence of transaction costs'', {\em Finance and Stochastics}, vol. 4, pp. 1-33. 
\bibitem{survey} Cadenilas A., (2000), ``Consumption-investment problems with transaction cost: survey and open problems,'', {\em Math. Methods Operations Research, vol 51}, pp,43-68. 
\bibitem{Dostal} Dostál P., ``Investment strategies in the long run with proportional transaction costs and a HARA utility function'',{\em Quantitative Finance, Vol. 9, No. 2, March 2009}.
\bibitem{MP} Ethier N., Kurtz G., (2005), ``Markov Processes'', John Wiley \& Sons, New Jersey. 
\bibitem{Portfolio} Follmer H., Schied A., (2004), ``Stochastic Finance. An Introduction in Discrete Time'', second edition, Walter de Gruyter, Berlin
\bibitem{Howard} Howard  R.A., (1972), ``Dynamic Programming and Markov Processes'', {\em Management Science}, Vol. 18, No. 7, March 1972.
\bibitem{HowardRS} Howard  R.A., Matheson J.E., (1960), ``Risk-Sensitive Markov Decision pocesses'', The M.I.T. Press, Princeton, NJ.
\bibitem{Howard} Howard R.A, Metheson J.E., (1972), ``Risk-Sensitive Markov Decision Processes'',. {\em Management Science}, Vol. 18, No. 7, March.   
\bibitem{Praskova} Karlin S., Taylor H.M., (1981), ``A Second Course in Stochastic Processes'', 
Academic Press, New York.
%\bibitem{Kal} Kalužíková M., ``Diskontované øízení portfolia'', Bachelor thesis {\em MFF UK}
\bibitem{KushnerDupuis} Kushner H.J., Dupuis P., (2001), ``Numerical Methods for Stochastic Control Problems in Continuous Time'',  Springer-Verlag, New York .  
\bibitem{Lancaster} Lancaster P., Tismenetsky M., (1985), ``The Theory of Matrices Second Edition'',  Academic Press.
\bibitem{Liggett} Liggett T.M., (2010), ``Continuous Time Markov Processes: An Introduction'',  American Mathematical Society, Providence.
\bibitem{MC} Norris J.R., (1997), ``Markov Chain'', Cambridge University Press, Cambridge. 
\bibitem{Merton} Merton R.C., (1992), ``Continuous-Time Finance'', B. Blackwell.
\bibitem{Shreve} Shreve S.E., (1981), ``Stochastic Calculus for Finance II'', Springer, New York.
\bibitem{Sta} Staníková D. (2006), ``Asymptotic Control of Portfolio'', Bachelor thesis, MFF UK.
\end{thebibliography}